---
title: "Meta-Analysis of "
author: "Dan Kerchner, Xinyang Zhang, Erika Hubbard, Xun Liu, Yijie He"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries and read study data:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(meta)

df <- read.csv('data/study_data.csv')
```

log-transform hazard ratios and compute standard error based on the confidence interval bounds 

```{r}
df <- df %>%
  mutate(logHR = log(HR),
         se_logHR = abs((log(HR_CI_upper) - log(HR_CI_lower))/(2*1.96)),
         trial = row_number(),
         study_year = paste(Study, Year)) %>%
  rename(`Hazard Ratio type` = HR_Estimated)

df$Quality <- factor(df$Quality, ordered = TRUE, levels = c('Poor', 'Fair', 'Good'))
```

# Perform meta-analysis using `metagen`

Based on https://link.springer.com/chapter/10.1007/978-3-319-21416-0_2

```{r fig.height=6, fig.width=10}
mg1 <- metagen(logHR, se_logHR,
               studlab=study_year, data = df,
               sm="HR", fixed = FALSE, random = TRUE)
print(mg1)
pdf(file='forestplot_main.pdf', height = 6, width = 10)
forest(mg1, sortvar = Year, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       print.tau2 = FALSE)
# didn't work:
#       leftcols = c("study_year", "TE", "seTE"),
#       leftlabs = c("Study (Year)", "log(HR)", "SE"))
dev.off()
forest(mg1, sortvar = Year, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       print.tau2 = FALSE)
```
# Subgroup analyses:

## Subgroup analysis by Hazard Ratio Type (Reported vs. Estimated)

```{r fig.height=6, fig.width=10}
mg2 <- update.meta(mg1, subgroup = `Hazard Ratio type`, control = list(maxiter=1000))
pdf(file='forestplot_HR_type.pdf', height = 6, width = 10)
forest(mg2, sortvar = Year, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       label.test.subgroup.common = 'Test for subgroup differences (fixed effects):',
       print.tau2 = FALSE)
dev.off()
forest(mg2, sortvar = Year, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       label.test.subgroup.common = 'Test for subgroup differences (fixed effects):',
       print.tau2 = FALSE)
```

## Subgroup analysis by Cancer Type

```{r fig.height=6, fig.width=10}
mg3 <- update.meta(mg1, subgroup = Cancer_Type)
pdf(file='forestplot_cancer_type.pdf', height = 6, width = 10)
forest(mg3, sortvar = Year, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       label.test.subgroup.common = 'Test for subgroup differences (fixed effects):',
       print.tau2 = FALSE)
dev.off()
forest(mg3, sortvar = Year, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       label.test.subgroup.common = 'Test for subgroup differences (fixed effects):',
       print.tau2 = FALSE)
```


## Subgroup analysis by treatment type

```{r fig.height=6, fig.width=10}
mg4 <- update.meta(mg1, subgroup = Treatment_Type)
pdf(file='forestplot_treatment_type.pdf', height = 7, width = 10)
forest(mg4, sortvar = Year, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       label.test.subgroup.common = 'Test for subgroup differences (fixed effects):',
       print.tau2 = FALSE)
dev.off()
forest(mg4, sortvar = Year, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       label.test.subgroup.common = 'Test for subgroup differences (fixed effects):',
       print.tau2 = FALSE)
```
## Subgroup analysis by study quality

```{r fig.height=7, fig.width=10}
mg5 <- update.meta(mg1, subgroup = Quality)
pdf(file='forestplot_quality.pdf', height = 7, width = 10)
forest(mg5, sortvar = Year, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       label.test.subgroup.common = 'Test for subgroup differences (fixed effects):',
       print.tau2 = FALSE)
dev.off()
forest(mg5, sortvar = Year, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       label.test.subgroup.common = 'Test for subgroup differences (fixed effects):',
       print.tau2 = FALSE)
```

# Publication Bias testing

## Funnel plot to visualize possible publication bias

```{r}
pdf(file='funnelplot.pdf', height = 6.5, width = 8.5)
funnel(mg1, studlab = df$study, cex.studlab = 0.5)#, xlim = c(0.55, 1.50))
dev.off()
funnel(mg1, studlab = df$study, cex.studlab = 0.5)#, xlim = c(0.55, 1.50))
```

## Begg's test of funnel plot asymmetry

```{r}
metabias(mg1, k.min=9, method.bias = "Begg")
```

## Egger's Linear regression test of funnel plot asymmetry

```{r}
metabias(mg1, k.min=9, plotit = TRUE, method.bias = "Egger")
```
# Sensitivity Analysis

## Leave-one-out analysis

```{r}
l1o <- metainf(mg1)
```

```{r}
pdf(file='leave1out.pdf', height = 6, width = 10)
forest(l1o, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       label.test.subgroup.common = 'Test for subgroup differences (fixed effects):',
       print.tau2 = FALSE)
dev.off()
forest(l1o, layout="RevMan5",
       text.common = 'Total (fixed effect, 95% CI)', text.w.common = 'fixed',
       label.left = "favors treatment", label.right = "favors control",
       label.test.subgroup.common = 'Test for subgroup differences (fixed effects):',
       print.tau2 = FALSE)
```

